Boosting the Accuracy of Finite Difference Schemes via Optimal Time Step Selection and Non-Iterative Defect Correction

### 2011-07-27 papers/OTSPDE/OTSPDE.tex
Ran OTSPDE manuscript through spell checker.

In Figure~\ref{fig:4th_order_parabolic_eqn_1d_perf}, observe that the 
2nd-order Crank-Nicholson scheme actually outperforms the OTS-NIDC scheme over 
the high (but still practically useful) range of errors even though 
Table~\ref{tab:comp_perf_vs_err_4th_order_parabolic} predicts that the OTS-NIDC
scheme should do better asymptotically.  The origin of this discrepancy is the
constants hidden by the big-$O$ notation.  In particular, the factor of $7/120$
in the optimal time step for the forward Euler scheme with a fourth-order 
approximation of the bilaplacian leads to a large hidden constant 
for the compute time for the OTS-NIDC scheme, and the low cost of inverting
the banded-diagonal biaplacian matrix in 1D leads to a relatively small hidden
constant in the Crank-Nicholson schemes.

As Figures~\ref{fig:4th_order_parabolic_eqn_1d_perf} and
\ref{fig:4th_order_parabolic_eqn_1d_error_roundoff} show, there is an
important advantage to explicit time integration schemes for the fourth-order
parabolic equation -- round-off error associated with matrix inversion does 
\emph{not} limit the achievable accuracy of the numerical solution.  In 
constrast, errors introduced in matrix inversion affects both the 2nd- and
4th-order Crank-Nicholson schemes.

@@

\bibitem{kress_2006}
W.~Kress.
\newblock Error estimates for deferred correction methods in time.
\newblock {\em Apll. Numer. Math}, 57:335--353, 2006.

>>>>

In Figure~\ref{fig:4th_order_parabolic_eqn_1d_perf}, observe that the 
2nd-order Crank-Nicholson scheme actually outperforms the OTS-NIDC scheme over 
the high (but still practically useful) range of errors even though 
Table~\ref{tab:comp_perf_vs_err_4th_order_parabolic} predicts that the OTS-NIDC
scheme should do better asymptotically.  The origin of this discrepancy is the
constants hidden by the big-$O$ notation.  In particular, the factor of $7/120$
in the optimal time step for the forward Euler scheme with a fourth-order 
approximation of the bilaplacian leads to a large hidden constant 
for the compute time for the OTS-NIDC scheme, and the low cost of inverting
the banded-diagonal bilaplacian matrix in 1D leads to a relatively small hidden
constant in the Crank-Nicholson schemes.

As Figures~\ref{fig:4th_order_parabolic_eqn_1d_perf} and
\ref{fig:4th_order_parabolic_eqn_1d_error_roundoff} show, there is an
important advantage to explicit time integration schemes for the fourth-order
parabolic equation -- round-off error associated with matrix inversion does 
\emph{not} limit the achievable accuracy of the numerical solution.  In 
contrast, errors introduced in matrix inversion affects both the 2nd- and
4th-order Crank-Nicholson schemes.

@@

\bibitem{kress_2006}
W.~Kress.
\newblock Error estimates for deferred correction methods in time.
\newblock {\em Appl. Numer. Math}, 57:335--353, 2006.
